---
title: "Asymptotic CI Validation: Simulation Results"
description: >
  Simulation-based calibration and coverage evaluation for all asymptotic
  confidence interval methods (normal, QF, and Cornish-Fisher) and the
  bootstrap reference, for linear and logistic models with both parametric
  and robust variance estimators.
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: flatly
---

<!--
AI INSTRUCTIONS FOR FUTURE SESSIONS
====================================
This document is the simulation-validation article for the RESI package website.
It lives at RESI/vignettes/articles/ci_validation.Rmd and is registered in
RESI/_pkgdown.yml under the "articles:" section.

PRE-COMPUTED FIGURES
All 36 PDFs referenced below are committed to
  RESI/vignettes/articles/figures/
and were produced by the simulation pipeline described in the
"Reproducing these figures" section of this document.

TO REGENERATE FIGURES after changing the simulation code or the estimator:
  1. Run the `eval=FALSE` code chunks in this document from the project root.
  2. Copy the output PDFs back to vignettes/articles/figures/ (the code chunk
     already does this as the last step).
  3. Commit the updated PDFs.

FIGURE SOURCES
  calibration_*.pdf      ← simCalibrationFigures(..., deriv_method="extended")
  compare_*.pdf          ← simCompareMethodsFigures(figures.dir="method_comparison")
  covquant_*.pdf         ← simCompareMethodsFigures(figures.dir="method_comparison")
  estimator_*.pdf        ← simCompareMethodsFigures(figures.dir="method_comparison")
  estimator_compare_*.pdf← simEstimatorFigures(figures.dir="method_comparison")
  sim_*_qf.pdf           ← simFigures(output.dir="resiAsympQFSim")

PKGDOWN
The article is listed under navbar articles in RESI/_pkgdown.yml.
vignettes/articles/ is excluded from the CRAN tarball via .Rbuildignore.

VARIANCE ESTIMATOR
All results use the extended influence-function Sigma_R estimator
(.resi_contrast_ext(), deriv_method="extended"). If the estimator changes,
regenerate all figures and update the exposition in section 2.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      out.width = "100%")
```

# Overview

This article presents simulation-based validation of the asymptotic confidence
interval (CI) methods available in the RESI package via `resi_pe_asymptotic()`
and `resi()`.

All asymptotic methods share the same $\hat\Sigma_R$ estimator — the
**extended influence-function** approach that treats the sandwich bread
$\hat\Sigma_{XA}$ and meat $\hat\Sigma_{XB}$ as independent estimators
(argument `deriv_method = "extended"` internally) — and differ only in how
they invert the asymptotic distribution of the Wald statistic
$T^2 = n\tilde{S}^2$.

| `ci.method` | Description |
|---|---|
| `"qf"` **(default)** | Quadratic-form CI: inverts the exact weighted chi-square distribution via the Imhof/Davies algorithm |
| `"normal"` | Truncated-normal CI using a bias-corrected normal approximation |
| `"cf"` | Cornish-Fisher CI: inverts a Cornish-Fisher expansion of the weighted chi-square |

Results are shown for four estimation settings:

| Setting | Model | Variance estimator |
|---|---|---|
| lm parametric | `lm()` | `stats::vcov` |
| lm robust | `lm()` | `sandwich::vcovHC` (HC3) |
| glm parametric | `glm(..., family=binomial)` | `stats::vcov` |
| glm robust | `glm(..., family=binomial)` | `sandwich::vcovHC` (HC3) |

For background on the RESI, see the
[RESI package website](https://statimagcoll.github.io/RESI/) and
[Jones et al. (2025)](https://doi.org/10.18637/jss.v112.i03).

---

# Simulation design

## Plasmode simulation

All evaluations use a **plasmode simulation design**
([Gadbury et al. 2008](https://doi.org/10.1177/0962280207080302))
built around the `insurance` dataset ($N = 1338$) shipped with the package.
Bootstrap resampling (with replacement) from the full dataset gives the
empirical sampling distribution.  RESI values estimated on the full dataset
via `resi_pe()` serve as the population targets.

Two models are evaluated:

- **Linear model** (`lm`):
  `log10(charges) ~ ns(age, df=3) * sex + bmi + smoker + region`
- **Logistic model** (`glm`, binomial):
  `I(charges > 15000) ~ ns(age, df=3) * sex + bmi + smoker + region`

Each simulation cell (model × variance estimator × sample size) uses
**1 000 replicates**. Results are shown for sample sizes
$n \in \{50, 100, 200, 500, 1000, 2000, 5000\}$.

## Metrics

For each model coefficient and ANOVA term the following quantities are reported:

- **Coverage**: proportion of replicates in which the 95% CI contains the
  population RESI value.
- **Width**: median CI width.
- **Bias** and **MSE**: mean signed and mean squared deviation of $\hat{S}$
  from the population value.

Bootstrap CIs (1 000 replicates, non-parametric) provide the reference standard.

## Variance calibration

The calibration study examines whether the analytic $\hat\sigma^2_R$ (the CI
width estimator) is consistent for the true Monte Carlo variance of $\hat{R}$
across replications.  Each calibration figure has four panels:

1. **Bias of $\hat\theta$**: signed deviation of the parameter estimates
   (model coefficients and, for `lm`, $\hat\phi = \hat\sigma^2$) from their
   full-data targets.
2. **Sandwich covariance normalized bias**: $(n\,\widehat{\mathrm{Var}}(\hat\beta_k) -
   n\,\mathrm{Var}_{\mathrm{MC}}(\hat\beta_k)) / n\,\mathrm{Var}_{\mathrm{MC}}(\hat\beta_k)$.
3. **Bias of $\hat{R}$** (RESI point estimate).
4. **$\sigma^2_R$ ratios**: solid line = $\sigma^2_{R,\text{MC}} /
   \sigma^2_{R,\text{extended}}$ (Monte Carlo variance over extended
   estimator); dashed = ratio against the initial delta-method derivation.
   A ratio of 1 (dashed reference line) indicates perfect calibration.

---

# How to reproduce these figures

The chunks below are **not evaluated** when the article is rendered.
Run them from a working directory of your choice; then copy the resulting
PDFs into `vignettes/articles/figures/` inside the package source tree.

## Step 1 — run plasmode simulations

```{r run-sims, eval=FALSE, echo=TRUE}
library(RESI)

# Bootstrap reference  (~hours; adjust mc.cores to your hardware)
insurancePlasmodeSim(nsim = 1000, ci.method = "boot",
                     output.dir = "resiBootSim",
                     mc.cores.settings = 4)

# Normal approximation
insurancePlasmodeSim(nsim = 1000, ci.method = "normal",
                     output.dir = "resiAsympNormalSim",
                     mc.cores.settings = 4)

# Quadratic-form (default)
insurancePlasmodeSim(nsim = 1000, ci.method = "qf",
                     output.dir = "resiAsympQFSim",
                     mc.cores.settings = 4)

# Cornish-Fisher
insurancePlasmodeSim(nsim = 1000, ci.method = "cf",
                     output.dir = "resiAsympCFSim",
                     mc.cores.settings = 4)
```

## Step 2 — run variance calibration simulation

```{r run-calib, eval=FALSE, echo=TRUE}
# 500 replicates per (model × vcov × n) cell; extended estimator
simCalibrationSim(nsim = 500,
                  output.dir   = "resiCalibrationSim",
                  deriv_method = "extended",
                  mc.cores.reps = 4)
```

## Step 3 — generate figures

```{r gen-figs, eval=FALSE, echo=TRUE}
# QF detail figures (one PDF per model/vcov setting)
simFigures(output.dir = "resiAsympQFSim")

# Variance calibration figures
simCalibrationFigures(output.dir  = "resiCalibrationSim",
                      deriv_method = "extended")

# CI method comparison, coverage quantile, and estimator bias/MSE figures
simCompareMethodsFigures(
  output.dirs = c(Bootstrap = "resiBootSim",
                  Normal    = "resiAsympNormalSim",
                  QF        = "resiAsympQFSim",
                  CF        = "resiAsympCFSim"),
  figures.dir = "method_comparison"
)

# t2S/f2S vs z2S/chisq2S estimator comparison (lm only)
simEstimatorFigures(sim.dir     = "resiBootSim",
                    figures.dir = "method_comparison")
```

## Step 4 — copy figures into the package

```{r copy-figs, eval=FALSE, echo=TRUE}
# Adjust pkg_dir to the root of your RESI package source tree
pkg_dir  <- "path/to/RESI"
figs_dir <- file.path(pkg_dir, "vignettes", "articles", "figures")
dir.create(figs_dir, recursive = TRUE, showWarnings = FALSE)

file.copy(list.files("resiCalibrationSim/figures/extended",
                     pattern = "\\.pdf$", full.names = TRUE), figs_dir)
file.copy(list.files("resiAsympQFSim/figures",
                     pattern = "\\.pdf$", full.names = TRUE), figs_dir)
file.copy(list.files("method_comparison",
                     pattern = "\\.pdf$", full.names = TRUE), figs_dir)
```

---

# Variance estimator calibration {.tabset}

These figures show whether $\hat\sigma^2_R$ — the width estimator shared by
all three asymptotic CIs — is well-calibrated as a function of sample size.

## LM parametric

```{r calib-lm-param}
knitr::include_graphics("figures/calibration_lm_parametric.pdf")
```

## LM robust

```{r calib-lm-robust}
knitr::include_graphics("figures/calibration_lm_robust.pdf")
```

## GLM parametric

```{r calib-glm-param}
knitr::include_graphics("figures/calibration_glm_parametric.pdf")
```

## GLM robust

```{r calib-glm-robust}
knitr::include_graphics("figures/calibration_glm_robust.pdf")
```

---

# CI method comparison {.tabset}

Coverage (top) and width (bottom) across all four CI methods — Bootstrap,
Normal, QF, and Cornish-Fisher — as a function of sample size.
The nominal coverage is 95% (dashed reference line).
Each panel shows results for one regression term; terms are labelled with
their population RESI value.

## LM parametric — coefficients

```{r cmp-lm-param-coef}
knitr::include_graphics("figures/compare_lm_parametric_coefficients.pdf")
```

## LM parametric — ANOVA

```{r cmp-lm-param-anova}
knitr::include_graphics("figures/compare_lm_parametric_anova.pdf")
```

## LM robust — coefficients

```{r cmp-lm-rob-coef}
knitr::include_graphics("figures/compare_lm_robust_coefficients.pdf")
```

## LM robust — ANOVA

```{r cmp-lm-rob-anova}
knitr::include_graphics("figures/compare_lm_robust_anova.pdf")
```

## GLM parametric — coefficients

```{r cmp-glm-param-coef}
knitr::include_graphics("figures/compare_glm_parametric_coefficients.pdf")
```

## GLM parametric — ANOVA

```{r cmp-glm-param-anova}
knitr::include_graphics("figures/compare_glm_parametric_anova.pdf")
```

## GLM robust — coefficients

```{r cmp-glm-rob-coef}
knitr::include_graphics("figures/compare_glm_robust_coefficients.pdf")
```

## GLM robust — ANOVA

```{r cmp-glm-rob-anova}
knitr::include_graphics("figures/compare_glm_robust_anova.pdf")
```

---

# Coverage quantile comparison {.tabset}

These figures show the distribution of per-term coverage across the sample-size
grid, displayed as quantile bands.  Unlike section 4, which shows mean coverage
per term, this view reveals how variable coverage is across terms and highlights
any terms where a method systematically under- or over-covers.

## LM parametric — coefficients

```{r covq-lm-param-coef}
knitr::include_graphics("figures/covquant_lm_parametric_coefficients.pdf")
```

## LM parametric — ANOVA

```{r covq-lm-param-anova}
knitr::include_graphics("figures/covquant_lm_parametric_anova.pdf")
```

## LM robust — coefficients

```{r covq-lm-rob-coef}
knitr::include_graphics("figures/covquant_lm_robust_coefficients.pdf")
```

## LM robust — ANOVA

```{r covq-lm-rob-anova}
knitr::include_graphics("figures/covquant_lm_robust_anova.pdf")
```

## GLM parametric — coefficients

```{r covq-glm-param-coef}
knitr::include_graphics("figures/covquant_glm_parametric_coefficients.pdf")
```

## GLM parametric — ANOVA

```{r covq-glm-param-anova}
knitr::include_graphics("figures/covquant_glm_parametric_anova.pdf")
```

## GLM robust — coefficients

```{r covq-glm-rob-coef}
knitr::include_graphics("figures/covquant_glm_robust_coefficients.pdf")
```

## GLM robust — ANOVA

```{r covq-glm-rob-anova}
knitr::include_graphics("figures/covquant_glm_robust_anova.pdf")
```

---

# Point estimate bias and MSE {.tabset}

These figures show how the RESI **point estimate** behaves across CI methods
(bias and MSE per term) as a function of sample size.  Since all methods use
the same plug-in estimator $\hat{S}$, bias and MSE should be identical across
methods; any divergence reflects numerical differences in small samples.

## LM parametric — coefficients

```{r est-lm-param-coef}
knitr::include_graphics("figures/estimator_lm_parametric_coefficients.pdf")
```

## LM parametric — ANOVA

```{r est-lm-param-anova}
knitr::include_graphics("figures/estimator_lm_parametric_anova.pdf")
```

## LM robust — coefficients

```{r est-lm-rob-coef}
knitr::include_graphics("figures/estimator_lm_robust_coefficients.pdf")
```

## LM robust — ANOVA

```{r est-lm-rob-anova}
knitr::include_graphics("figures/estimator_lm_robust_anova.pdf")
```

## GLM parametric — coefficients

```{r est-glm-param-coef}
knitr::include_graphics("figures/estimator_glm_parametric_coefficients.pdf")
```

## GLM parametric — ANOVA

```{r est-glm-param-anova}
knitr::include_graphics("figures/estimator_glm_parametric_anova.pdf")
```

## GLM robust — coefficients

```{r est-glm-rob-coef}
knitr::include_graphics("figures/estimator_glm_robust_coefficients.pdf")
```

## GLM robust — ANOVA

```{r est-glm-rob-anova}
knitr::include_graphics("figures/estimator_glm_robust_anova.pdf")
```

---

# t/F vs z/$\chi^2$ estimator comparison (lm) {.tabset}

The `t2S`/`f2S` functions convert test statistics using a
**t- or F-distribution** quantile (accounting for residual degrees of
freedom); `z2S`/`chisq2S` use the **normal or chi-square** quantile.
These figures show how the two estimator families compare in bias and MSE
for the `lm` model, based on bootstrap simulation replicates.

## LM parametric — coefficients

```{r estcmp-lm-param-coef}
knitr::include_graphics("figures/estimator_compare_lm_parametric_coefficients.pdf")
```

## LM parametric — ANOVA

```{r estcmp-lm-param-anova}
knitr::include_graphics("figures/estimator_compare_lm_parametric_anova.pdf")
```

## LM robust — coefficients

```{r estcmp-lm-rob-coef}
knitr::include_graphics("figures/estimator_compare_lm_robust_coefficients.pdf")
```

## LM robust — ANOVA

```{r estcmp-lm-rob-anova}
knitr::include_graphics("figures/estimator_compare_lm_robust_anova.pdf")
```

---

# QF method detail {.tabset}

Per-term coverage and width for the quadratic-form (QF) CI — the package
default — across the full sample-size grid.

## LM parametric

```{r qf-lm-param}
knitr::include_graphics("figures/sim_lm_parametric_qf.pdf")
```

## LM robust

```{r qf-lm-rob}
knitr::include_graphics("figures/sim_lm_robust_qf.pdf")
```

## GLM parametric

```{r qf-glm-param}
knitr::include_graphics("figures/sim_glm_parametric_qf.pdf")
```

## GLM robust

```{r qf-glm-rob}
knitr::include_graphics("figures/sim_glm_robust_qf.pdf")
```

---

# Session info

```{r session}
sessionInfo()
```
